---
title: "remlTemplate"
author: "SRE"
date: Sys.Date()
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eml-2.1.1, include=FALSE}
options("emld_db" = "eml-2.1.1")
emld::eml_version("eml-2.2.0")
emld::eml_version("eml-2.1.1")
```

```{r libraries}
library(EML)
library(RPostgreSQL)
library(RMySQL)
library(tidyverse)
library(tools)
library(readxl)
library(aws.s3)
library(capeml)
library(gioseml)
```

```{r dataset_details}
projectid <- 
packageIdent <- 'knb-lter-cap.xxx.x'
pubDate <- as.character(Sys.Date())
```

```{r helper_functions}
source('~/localRepos/reml-helper-tools/amazon_file_upload.R')
```

```{r connections::amazon}
source('~/Documents/localSettings/aws.s3')
```

```{r connections::postgres::local, eval=FALSE}
source('~/Documents/localSettings/pg_local.R')
pg <- pg_local
```

```{r connections::postgres::prod, eval=T }
source('~/Documents/localSettings/pg_prod.R')
pg <- pg_prod
```

```{r connections::mysql::prod, eval=T }
source('~/Documents/localSettings/mysql_prod.R')
mysql_prod <- mysql_prod_connect()
```

```{r compiled_data, eval=TRUE}

# tabularDataEntity <- import / generate...process...

write_attributes(tabularDataEntity)
write_factors(tabularDataEntity)

tabularDataEntity_desc <- "desc"

tabularDataEntity_DT <- create_dataTable(dfname = tabularDataEntity,
                                              description = tabularDataEntity_desc)

```

```{r otherEntity, eval=TRUE}

otherEntityObject_OE <- create_otherEntity(targetFile = 'kml_pdf_etc',
                                  description = 'desc')

```

```{r spatialVector, eval=TRUE}

# call this library as needed
library(sf)

# note layer name is sans file extention
spatialVectorEntity <- read_sf(dsn = "directory/", layer = "layer name")

# assign Name to the column that is the site identifier
spatialVectorEntity <- spatialVectorEntity %>% 
  mutate(Name = site identifier)

write_attributes(spatialVectorEntity)

spatialVectorEntity_desc <- "desc"

spatialVector_SV <- create_spatialVector(targetFile = spatialVectorEntity,
                                         description = spatialVectorEntityDesc )

```

```{r title}

title <- ''
```

```{r abstract}

abstract <- set_TextType("abstract.md")
```

```{r connections::mysql::prod redux, eval=T }
source('~/Documents/localSettings/mysql_prod.R')
mysql_prod <- mysql_prod_connect()
```

```{r people}

# see gioseml for examples of creating people resources from scratch

mysql_prod <- mysql_prod_connect()

cameron <- create_role(firstName = "cameron", lastName = "boehme", roleType = "creator")
chelsea <- create_role(firstName = "chelsea", lastName = "stratton", roleType = "creator")
fabio <- create_role(giosPersonId = 23857, roleType = "creator")

creators <- list(cameron, chelsea, fabio)

cameron <- create_role(firstName = "cameron", lastName = "boehme", roleType = "metadata")
metadataProvider <- list(cameron)

```

```{r keywords}

# CAP IRTs for reference (be sure to include these as appropriate):
# https://sustainability.asu.edu/caplter/research/

write_keywords()
keywords <- create_keywordSet('keywords.csv')

```

```{r methods}

methods <- set_methods("methods.md")
```

```{r coverages}

# begindate <- format(min(runoff_chemistry$runoff_datetime), "%Y-%m-%d")
# enddate <- format(max(runoff_chemistry$runoff_datetime), "%Y-%m-%d")
geographicDescription <- "CAP LTER study area"
coverage <- set_coverage(begin = "2014-09-01",
                         end = "2015-03-30",
                         geographicDescription = geographicDescription,
                         west = -112.100, east = -111.877,
                         north = +33.608, south = +33.328)

```

New approach for taxonomy is to use taxonomyCleanr to build the
taxonomicCoverage. Note that at the time of this writing and building,
taxonomyCleanr had not been ported to rOpenSci EML v2. Using
caplter::taxonomyCleanr/taxonomy-rEML2 until Colin adapts taxonmyCleanr to
ropensci EML v2.

*Note* that the `taxa_map.csv` built with the `create_taxa_map()` function and
resolving taxonomic IDs (i.e., `resolve_comm_taxa()`) only needs to be run once
per version/session -- the taxonomicCoverage can be built as many times as
needed with `resolve_comm_taxa()` once the `taxa_map.csv` has been generated and
the taxonomic IDs resolved.

```{r taxonomyCleanr, eval=TRUE}

devtools::load_all('~/localRepos/taxonomyCleanr/')
library(taxonomyCleanr)

my_path <- getwd() # taxonomyCleanr requires a path (to build the taxa_map)

# plant taxa listed in the om_transpiration_factors file
plantTaxa <- read_csv('om_transpiration_factors.csv') %>% 
  filter(attributeName == "species") %>% 
  as.data.frame()

# create or update map. A taxa_map.csv is the heart of taxonomyCleanr. This
# function will build the taxa_map.csv and put it in the path identified with
# my_path.
create_taxa_map(path = my_path, x = plantTaxa, col = "definition") 

# resolve taxa by attempting to match the taxon's In this case, data.source 3 is
# ITIS (which, by the way, is the only authority taxonomyCleanr will allow for
# common names).
# resolve_comm_taxa(path = my_path, data.sources = 3) # in this case, 3 is ITIS
resolve_sci_taxa(path = my_path, data.sources = 3) # in this case, 3 is ITIS

# build the EML taxonomomic coverage
taxaCoverage <- make_taxonomicCoverage(path = my_path)

# add taxonomic to other coverages
coverage$taxonomicCoverage <- taxaCoverage
```

```{r construct-dataset}

# from capeml package:
# address
# publisher
# contact
# rights
# distribution

# DATASET
dataset <- EML::eml$dataset(
  title = title,
  creator = creators,
  pubDate = pubDate,
  metadataProvider = metadataProvider,
  intellectualRights = capRights,
  abstract = abstract,
  keywordSet = keywords,
  coverage = coverage,
  contact = capContact, # cap contact
  publisher = capPublisher, # cap pub
  # contact = giosContact, # gios contact
  # publisher = giosPublisher, # gios pub
  methods = methods,
  # project = capProject, # cap project
  distribution = create_distribution(packageIdent))

# add associatedParty if relevant
# dataset$associatedParty <- list() 

```

```{r dataSet$dataTable}

# add dataTables if relevant

print(ls(pattern = "_DT"))

if (length(ls(pattern = "_DT")) > 0) {
  
  listOfDataTables <- lapply(ls(pattern = "_DT"), function(DT) { get(DT) } )
  
  dataset$dataTable  <- listOfDataTables  
  
}

# or add manually
# dataset$dataTable <- list(dataTableOne, dataTableTwo)

```

```{r dataSet$otherEntity}

# add other entities if relevant

print(ls(pattern = "_OE"))

if (length(ls(pattern = "_OE")) > 0) {
  
  listOfOtherEntities <- lapply(ls(pattern = "_OE"), function(OE) { get(OE) } )
  
  dataset$otherEntity <- listOfOtherEntities 
  
}

# or add manually
# dataset$otherEntity <- list(otherEntityOne, otherEntityTwo)

```

```{r dataSet$spatialVector}

# add spatial vectors if relevant

print(ls(pattern = "_SV"))

if (length(ls(pattern = "_SV")) > 0) {
  
  listOfSpatialVectors <- lapply(ls(pattern = "_SV"), function(SV) { get(SV) } )
  
  dataset$spatialVector  <- listOfSpatialVectors  
  
}

# or add manually
# dataset$spatialVector <- list(spatialVectorOne, spatialVectorTwo)

```

```{r custom-units, eval=FALSE}

custom_units <- rbind(
  data.frame(id = "milligramPerKilogram",
             unitType = "massPerMass",
             parentSI = "gramsPerGram",
             multiplierToSI = 0.000001,
             description = "millgram of element per kilogram of material"))

unitList <- set_unitList(custom_units,
                         as_metadata = TRUE)

```

```{r literature cited, eval=TRUE}

# add literature cited if relevant

dataset$literatureCited <- citationsAll 

```

```{r construct_eml, eval=TRUE}

if(exists('unitList')) {
  
  eml <- EML::eml$eml(
    access = lterAccess,
    dataset = dataset,
    additionalMetadata = unitList,
    packageId = packageIdent,
    system = "knb",
    scope = "system"
  )
  
} else {
  
  eml <- EML::eml$eml(
    access = lterAccess,
    dataset = dataset,
    packageId = packageIdent,
    system = "knb",
    scope = "system"
  )
}

```

```{r write_eml}

# write the eml to file
write_eml(eml, paste0(packageIdent, ".xml"))
```

```{r preview_data_file_to_upload}

# preview data set files that will be uploaded to S3
list.files(pattern = paste0(projectid, "_"))
```

```{r upload_data_S3}

# upload files to S3
lapply(list.files(pattern = paste0(projectid, "_")), dataToAmz)
```

```{r clean_up}

# remove data files
dataFilesToRemove <- dir(pattern = paste0(projectid, "_"))
file.remove(dataFilesToRemove)

# EML to S3
if(length(list.files(pattern = "*.xml")) == 1) {
  emlToAmz(list.files(pattern = "*.xml")) } else {
    print("more than one xml file found")
  }

# EML to cap-data-eml and remove file from project
tryCatch({
  
  if(length(list.files(pattern = "*.xml")) == 1) {
    file.copy(list.files(pattern = "*.xml"), "/home/srearl/localRepos/cap-metadata/cap-data-eml/")
    file.remove(list.files(pattern = "*.xml")) } else {
      print("more than one xml file found")
    }
},
warning = function(warn) {
  print(paste("WARNING: ", warn))
},
error = function(err) {
  print(paste("ERROR: ", err))
  
}) # close try catch
```
