---
output: github_document
---

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}

library(kableExtra)
```

<!-- README.md is generated from README.Rmd. Please edit the latter. -->

## capeml: tools to aid the generation of EML metadata

### overview

This package contains tools to aid the generation of EML metadata with intent to
publish a data set (data + metadata) in the Environmental Data Initiative (EDI)
data repository. Functions and a workflow are included that allow for the
creation of metadata at the dataset level, and individual data entities (e.g.,
spatial vectors, raster, other entities, data tables).

Helper functions for the creation of dataset metadata, and `dataTable()`,
`otherEntity()`, `spatialRaster()`, and `spatialVector()` entities using the
[ropensci::eml package](https://ropensci.github.io/EML/) are supported.

Note that the creation of people-related entities in EML are specific to
functions that rely on Global Institute of Sustainability infrastructure; please
see the [gioseml](https://github.com/CAPLTER/gioseml) package for these tools.

A template workflow to generating a complete EML record is included with this
package:
[knb-lter-cap.xxx.Rmd](https://github.com/CAPLTER/capeml/blob/master/knb-lter-cap.xxx.Rmd).


### installation

Install from GitHub (after installing the
[devtools](https://cran.r-project.org/web/packages/devtools/index.html) package:

```{r load capeml package, eval=FALSE}

devtools::install_github("CAPLTER/capeml")
```

### options

#### EML

This package defaults to the current version of EML. If desired, users can
switch to the previous version with `emld::eml_version("eml-2.1.1")`.

#### project naming

Most EML-generating functions in the capeml package will create both physical
objects and EML references to those objects with the format:
`project-id`\_`object-name`\_`object-hash`\.`file-extension` (e.g.,
*664_site_map_5fb7b8d53d48010eab1a2e73db7f1941.png*). The target object (e.g.,
site_map.png from the previous example) is renamed with the additional metadata
and this object name is referenced in the EML metadata. The only exception to
this approach are spatialVectors where the hash of the file/object is not
included in the new object name. Note that the project-id is not passed to any
of the functions, and must exist in the working R environment (as `projectid`).

Project-naming functionality can be turned off by setting the `projectNaming`
option in `create_dataTable()`, `create_spatialRaster()`,
`create_spatialVector()`, and `create_otherEntity()` to FALSE. When set to
FALSE, the object name is not changed, and the file name of the object is
included in the EML.


### tools to generate entity metadata

**building dataTable, spatialRaster, and spatialVector metadata**

* `write_attributes()` creates a template as a csv file for supplying attribute
metadata for a tabular data object that resides in the R environment
* `write_factors()` creates a template as a csv file for supplying code
definition metadata for factors in a tabular data object that resides in the R
environment
* `write_raster_factors()` creates a template as a csv file for supplying code
definition metadata for spatial rasters if raster values are categorical

### tools to create EML entities

**constructing dataset elements**

* `create_dataTable()` creates a EML entity of type dataTable
* `create_otherEntity()` creates a EML entity of type otherEntity
* `create_spatialRaster()` creates a EML entity of type spatialRaster - see
[vignette](https://caplter.github.io/capeml/articles/create_spatialRaster.html)
for more detail
* `create_spatialVector()` creates a EML entity of type spatialRaster

### construct a dataset

#### abstract

The `create_dataset` function will look for a `abstract.md` file in the working
directory or at the path provided if specified. `abstract.md` must be a markdown
file.

#### keywords

`write_keywords()` creates a template as a csv file for supplying dataset
keywords. The `create_dataset` function will look for a `keywords.csv` file in
the working directory or at the path provided if specified.

#### methods

The `create_dataset` function will look for a `methods.md` file in the working
directory or at the path provided if specified (`methods.md` must be a markdown
file.). Alternatively, the workflow below is an enhanced approach of developing
methods if provenance data are required or there are multiple methods files.

```{r methods, eval=FALSE}

library(EDIutils)

# methods from file tagged as markdown
main <- read_markdown("~/Dropbox/development/knb-lter-cap.683/methods.md")

# provenance: naip
naip <- emld::as_emld(EDIutils::api_get_provenance_metadata("knb-lter-cap.623.1"))
naip$`@context` <- NULL
naip$`@type` <- NULL

# provenance: lst
lst <- emld::as_emld(EDIutils::api_get_provenance_metadata("knb-lter-cap.677.1"))
lst$`@context` <- NULL
lst$`@type` <- NULL

enhancedMethods <- EML::eml$methods(methodStep = list(main, naip, lst))

```

#### coverages

*Geographic* and *temporal* coverages are straight foward and
documented in the workflow, but creating a *taxonomic* coverage is more
involved. *Taxonomic coverage(s)* are constructed using EDI's
[taxonomyCleanr](https://github.com/EDIorg/taxonomyCleanr) tool suite.

A sample workflow for creating a taxonomic coverage:

```{r taxonomyCleanr, eval = FALSE}

library(taxonomyCleanr)

my_path <- getwd() # taxonomyCleanr requires a path (to build the taxa_map)

# Example: draw taxonomic information from existing resource:

# plant taxa listed in the om_transpiration_factors file
plantTaxa <- read_csv('om_transpiration_factors.csv') %>% 
  filter(attributeName == "species") %>% 
  as.data.frame()

# create or update map. A taxa_map.csv is the heart of taxonomyCleanr. This
# function will build the taxa_map.csv and put it in the path identified with
# my_path.
create_taxa_map(path = my_path, x = plantTaxa, col = "definition") 

# Example: construct taxonomic resource:

gambelQuail <- tibble(taxName = "Callipepla gambelii")

# Create or update map: a taxa_map.csv is the heart of taxonomyCleanr. This
# function will build the taxa_map.csv in the path identified with my_path.
create_taxa_map(path = my_path, x = gambelQuail, col = "taxName") 

# Resolve taxa by attempting to match the taxon name (data.source 3 is ITIS but
# other sources are accessible). Use `resolve_comm_taxa` instead of
# `resolve_sci_taxa` if taxa names are common names but note that ITIS
# (data.source 3) is the only authority taxonomyCleanr will allow for common
# names.
resolve_sci_taxa(path = my_path, data.sources = 3) # in this case, 3 is ITIS

# build the EML taxonomomic coverage
taxaCoverage <- make_taxonomicCoverage(path = my_path)

# add taxonomic to the other coverages
coverage$taxonomicCoverage <- taxaCoverage
```

### overview: create a dataTable

Given a rectangular data matrix of type dataframe or Tibble in the R
environment:

`write_attributes(data_entity)` will generate a template as a csv file in the
working directory based on properties of the data entity such that metadata
properties (e.g., attributeDefinition, units) can be added via a editor or
spreadsheet application.

`write_factors(data_entity)` will generate a template as a csv file in the
working directory based on columns of the data entity that are factors such that
details of factor levels can be added via a editor or spreadsheet application.

`create_dataTable(data_entity)` performs many services:

* the data entity is written to file as a csv in the working directory with the
file name: *projectid_data-entity-name_md5-hash-of-file.csv*
* metadata provided in the attributes and factors (if relevant) templates are
ingested
* a EML object of type dataTable is returned
* note that the data entity name should be used consistently within the chunk,
and the resulting dataTable entity should have the name: *data_entity_DT*

### overview: create a otherEntity

A EML object of type otherEntity can be created from a single file or a
directory. In the case of generating a otherEntity object from a directory, pass
the directory path to the targetFile argument, capeml will recognize the target
as a directory, and create a zipped file of the identified directory.

If the otherEntity object already is a zip file with the desired name, set the
overwrite argument to FALSE to prevent overwriting the existing object.

As with all objects created with the capeml package, the resulting object is
named with convention: projectid_object-name_md5sum-hash.file extension by
default but this functionality can be turned off by setting projectNaming to
FALSE.

### literature cited

Though not provided as a function, below is a workflow for adding literature
cited elements at the dataset level. The workflow capitalizes on EML version 2.2
that accepts the BibTex format for references.

```{r dataset-level-citations, eval=FALSE}

# add literature cited if relevant
library(rcrossref)
library(EML)

mccafferty <- cr_cn(dois = "https://doi.org/10.1186/s40317-015-0075-2", format = "bibtex")
mccaffertycit <- EML::eml$citation(id = "https://doi.org/10.1186/s40317-015-0075-2")
mccaffertycit$bibtex <- mccafferty 

citations <- list(
  citation = list(
    mccaffertycit
  ) # close list of citations
) # close citation

dataset$literatureCited <- citations 

```
