---
output: github_document
---

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}

library(kableExtra)
```

<!-- README.md is generated from README.Rmd. Please edit the latter. -->

# capeml: tools to aid the generation of EML metadata

### overview

This package contains tools to aid the generation of EML metadata with intent to publish a data set (data + metadata) in the Environmental Data Initiative (EDI) data repository. Functions and a workflow are included that allow for the creation of metadata at the data set level, and individual data entities (e.g., spatial vectors, raster, other entities, data tables). 

Helper functions for the creation of data set metadata and a dataTable entity using the ropensci/eml are currently supported. Helper functions for the creation of metadata for spatial data (e.g., spatialRaster) currently are built on an earlier version of ropensci/eml and not functional - these will be updated soon (2018-12-17). Note that the creation of people-related entities in EML are specific to functions that rely on Global Institute of Sustainability infrastructure - these tools are located in a the [gioseml](https://github.com/CAPLTER/gioseml) package.

**navigation**

- [installation](https://github.com/CAPLTER/capeml#installation)
- [dataTable](https://github.com/CAPLTER/capeml#dataTable)
- [dataSet](https://github.com/CAPLTER/capeml#dataSet)

currently unsupported
- [spatial raster](https://github.com/CAPLTER/capeml#generate-spatialraster)
- [generate keyword set](https://github.com/CAPLTER/capeml#generate-keywordset-from-file)
- [generate taxonomic coverage](https://github.com/CAPLTER/capeml#generate-taxonomiccoverage)


### installation

Install the current version from GitHub (after installing the `devtools` package from CRAN):

```{r load package, eval=FALSE}

devtools::install_github("CAPLTER/capeml")
```


### dataTable

Given a rectangular data matrix of type dataframe or Tibble in the R envionment:

`write_attributes(data_entity)` will generate a template as a csv file in the working directory based on properties of the data entity such that metadata properties (e.g., attributeDefinition, units) can be added via a editor or spreadsheet application. 

`write_factors(data_entity)` will generate a template as a csv file in the working directory based on columns of the data entity that are factors such that details of factor levels can be added via a editor or spreadsheet application.

`create_dataTable(data_entity)` performs many services:
* the data entity is written to file as a csv in the working directory with the file name: projectid_data-entity-name_md5-hash-of-file.csv
* metadata provided in the attributes and factors (if relevant) templates are ingested
* a EML object of type dataTable is returned
* note that the data entity name should be used consistently within the chunk, and the resulting dataTable entity should have the name: data_entity_DT

```{r sample_createDataTable, eval=FALSE}

data_entity <- read("data source") %>% 
  processing %>% 
  processing

write_attributes(data_entity)
write_factors(data_entity)

data_entity_desc <- "snow leopard data"

data_entity_DT <- create_dataTable(dfname = data_entity,
                                   description = data_entity_desc)

```


### dataSet

...most documentation in project template (in progress)...

`write_keywords()` will generate a template as a csv file in the working directory with default CAP LTER keywords, and the appropriate thesaurii for adding and referencing additional keywords.


### generate spatialRaster

Please see the [create_spatialRaster](https://github.com/CAPLTER/capeml/blob/master/vignettes/create_spatialRaster.Rmd) vignette for more information about creating EML for raster data.


### generate keywordSet (from file)

The function `create_keywordSet` generates a EML object of type keywordSet from a csv file containing thesaurus, keyword, and type where type is an optional keyword attribute. Keyword files should be structured like the following...

```{r sampleKeywordFile, echo=FALSE}

load(file = './data/keywordTest.rda')
kable(keywordTest)
```

Call the function:

```{r create_keywordSet, eval=FALSE}

datasetKeywords <- create_keywordSet('path/keywordFile.csv')
```

The object (e.g., datasetKeywords) can then be included in the EML generation script.

### generate taxonomicCoverage

The function `set_taxonomicCoverage` is borrowed from [rOpenSci EML](https://github.com/ropensci/EML) to create a EML entity of type taxonomicCoverage that can be added to the coverage element of an EML document. This function and approach is a fast-and-easy way to generate a taxonomicCoverage for resolvable taxa in a data set - it is not designed to resolve or address any challenges to resolving the ITIS identification of a taxa. That is, only taxa that are cleanly matched to the ITIS database are included in the taxonomicCoverage. As such, some taxa may not be included in the taxonomicCoverage element. For a more iterative approach that allows for resolving the corresponding ITIS ID of all taxa (but also may require taxonomic expertise), users should consider the EDI's [taxonomyCleanr](https://github.com/EDIorg/taxonomyCleanr) package. 

Taxa should first be passed through the `indentify_resolvable_taxa` function to identify those taxa for which an ITIS ID is resolvable.

```{r identify_resolvable_taxa, eval=FALSE}

all_taxa <- identify_resolvable_taxa(listOfTaxa)
```

Run setTaxonomicCoverage on the subset of taxa for which an ITIS ID is resolvable.

```{r set_taxonomicCoverage workflow, eval=FALSE}

resolved_taxa <- c(all_taxa[!is.na(all_taxa$resolve),]$taxon)a
taxaCoverage <- set_taxonomicCoverage(resolved_taxa, expand = T, db = 'itis')
```

```{r add taxonomic coverage to EML coverage, eval=FALSE}

coverage@taxonomicCoverage <- c(taxaCoverage)
```


### generate otherEntity with dataType kml

The function `create_KML`...

```{r generate otherEntity of dataType kml, eval=FALSE}

desert_fertilization_sites <- create_KML(
  kmlFile = "~/Desktop/desert_fertilization_sampling_sites.kml",
  description = "approximate location of desert fertiliztion long-term study sites")
```
