---
output: github_document
---

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}

library(kableExtra)
```

<!-- README.md is generated from README.Rmd. Please edit the latter. -->

## capeml: tools to aid the generation of EML metadata

### overview

This package contains tools to aid the generation of EML metadata with intent to
publish a data set (data + metadata) in the Environmental Data Initiative (EDI)
data repository. Functions and a workflow are included that allow for the
creation of metadata at the data set level, and individual data entities (e.g.,
spatial vectors, raster, other entities, data tables).

Helper functions for the creation of data set metadata and a dataTable entity
using the ropensci/eml are currently supported.

Note that the creation of people-related entities in EML are specific to
functions that rely on Global Institute of Sustainability infrastructure; please
see the [gioseml](https://github.com/CAPLTER/gioseml) package for these tools.

**navigation**

* [installation](https://github.com/CAPLTER/capeml#installation)
* [project-naming](https://github.com/CAPLTER/capeml#project-naming)
* [dataSet element: keywords, taxonomic coverage](https://github.com/CAPLTER/capeml#keywords)
* [dataSet element: taxonomic coverage](https://github.com/CAPLTER/capeml#taxonomic coverage)
* [dataTable](https://github.com/CAPLTER/capeml#dataTable)
* [otherEntity](https://github.com/CAPLTER/capeml#otherEntity)
* [spatialVector](https://github.com/CAPLTER/capeml#spatialVector)
- [spatialRaster](https://github.com/CAPLTER/capeml#spatialraster)


### installation

Install the current version from GitHub (after installing the `devtools` package from CRAN):

```{r load capeml package, eval=FALSE}

devtools::install_github("CAPLTER/capeml")
```


### project-naming

Most EML-generating functions in the capeml package will create both physical
objects and EML references to those objects with the format:
`project-id`_`object-name`_`object-hash`.`file-extension` (e.g.,
664_site_map_5fb7b8d53d48010eab1a2e73db7f1941.png). The target object (e.g.,
site_map.png from the previous example) is renamed with the additional metadata
and this object name is referenced in the EML metadata. The only exception to
this approach are spatialVectors where the hash of the file/object is not
included in the new object name. Note that the project-id is not passed to any
of the functions, and must exist in the working R environment (as `projectid`).

project-naming functionality can be turned off by setting the `projectNaming`
option in `create_dataTable`, `create_spatialRaster`, `create_spatialVector`,
and `create_otherEntity` to FALSE. When set to FALSE, the object name is not
changed, and the file name of the object is included in the EML.


### dataSet elements: keywords, taxonomic coverage

#### keywords

`write_keywords()` will generate a template as a csv file titled `keywords.csv`
in the working directory with default CAP LTER keywords, and the appropriate
thesaurii for adding and referencing additional keywords.

Once constructed, keywords can be added to the dataSet by referencing the
keyword file with ropensci/EML::create_keywordSet

```{r create keyword set, eval=FALSE}
keywords <- create_keywordSet('keywords.csv')
```

#### taxonomic coverage

New approach for taxonomy is to use EDI's taxonomyCleanr to build the
taxonomicCoverage. Note that at the time of this writing (2019-05-03),
taxonomyCleanr had not been ported to rOpenSci EML v2. As such, this workflow
employs a forked, feature branch of
[taxonomyCleanr](https://github.com/CAPLTER/taxonomyCleanr/tree/taxonomy-rEML2)
until EDI adapts taxonmyCleanr to ropensci EML v2.

*Note* that the `taxa_map.csv` built with the `create_taxa_map()` function and
resolving taxonomic IDs (i.e., `resolve_comm_taxa()`) only needs to be run once
per version/session -- the taxonomicCoverage can be built as many times as
needed with `resolve_comm_taxa()` or `resolve_sci_taxa` once the `taxa_map.csv`
has been generated and the taxonomic IDs resolved.

A sample workflow:

```{r create taxonomic coverage, eval=FALSE}

# load and call the feature branch of taxonomyCleanr
devtools::load_all('path-to-repo/taxonomyCleanr/') 
library(taxonomyCleanr)

# taxonomyCleanr requires a path to build the taxa_map
my_path <- getwd() 

# taxonomic data must be in a data frame (load if not already in the environment)
myTaxa <- read_csv('file with taxa') %>% 
  as.data.frame()

# create or update map. A taxa_map.csv is the heart of taxonomyCleanr. This
# function will build the taxa_map.csv and put it in the path identified with
# my_path.
create_taxa_map(path = my_path,
                x = myTaxa,
                col = "myTaxa column with taxonomic names") 

# resolve taxa by attempting to match the taxon names. In this case, data.source
# 3 is ITIS (which, as of 2019-05-03, is the only authority taxonomyCleanr will
# allow for common names).

# resolve from common name:
resolve_comm_taxa(path = my_path, data.sources = 3) # in this case, 3 is ITIS

-- OR --

# resolve from scientific name:
resolve_sci_taxa(path = my_path, data.sources = 3) # in this case, 3 is ITIS

# build the EML taxonomomic coverage
taxaCoverage <- make_taxonomicCoverage(path = my_path)

# add taxonomic to other coverages
coverage$taxonomicCoverage <- taxaCoverage
```


### dataTable

Given a rectangular data matrix of type dataframe or Tibble in the R envionment:

`write_attributes(data_entity)` will generate a template as a csv file in the
working directory based on properties of the data entity such that metadata
properties (e.g., attributeDefinition, units) can be added via a editor or
spreadsheet application.

`write_factors(data_entity)` will generate a template as a csv file in the
working directory based on columns of the data entity that are factors such that
details of factor levels can be added via a editor or spreadsheet application.

`create_dataTable(data_entity)` performs many services:

* the data entity is written to file as a csv in the working directory with the
file name: *projectid_data-entity-name_md5-hash-of-file.csv*
* metadata provided in the attributes and factors (if relevant) templates are
ingested
* a EML object of type dataTable is returned
* note that the data entity name should be used consistently within the chunk,
and the resulting dataTable entity should have the name: *data_entity_DT*

The `create_dataTable` function accepts five arguments:

1. dfname = dataframe or tibble object name (required)
2. description = a description of the data object (required)
3. dateRangeField = (optional) The quoted name of the data entity field that is a
date field that would reflect the start and end dates of the data reflected in
the data entity. The date must be of the format %Y-%m-%d.
4. baseURL = The base path of the web-accessible location of the data
file; the name of the resulting file will be passed to the base path to generate
a web-resolvable file path.
5. missingValueCode = (optional) create_dataTable will automatically document
the presence of NA and NaN entries as missing values in the EML output. The user
has the ability to identify an additional indicator of missing values. Numbers
(e.g., -9999 should be unquoted) whereas text values (e.g., "missing") should be
quoted.

```{r sample_createDataTable, eval=FALSE}

data_entity <- read("data source") %>% 
  processing %>% 
  processing

write_attributes(data_entity)
write_factors(data_entity)

data_entity_desc <- "snow leopard data"

# create_dataTable with minimal arguments
data_entity_DT <- create_dataTable(dfname = data_entity,
                                   description = data_entity_desc)

# create_dataTable with optional arguments (sans baseURL)
data_entity_DT <- create_dataTable(dfname = data_entity,
                                   description = data_entity_desc,
                                   dateRangeField = "observation date",
                                   missingValueCode = "missing")

```


### otherEntity

A EML object of type otherEntity can be created from a single file or a
directory. In the case of generating a otherEntity object from a directory, pass
the directory path to the targetFile argument, capeml will recognize the target
as a directory, and create a zipped file of the identified directory.

If the otherEntity object already is a zip file with the desired name, set the
overwrite argument to FALSE to prevent overwriting the existing object.

As with all objects created with the capeml package, the resulting object is
named with convention: projectid_object-name_md5sum-hash.file extension by
default but this functionality can be turned off by setting projectNaming to
FALSE.

```{r sample_createOtherEntity, eval=FALSE}

# general use
other_entity_name <- create_otherEntity(
  targetFile = "file or directory name",
  description = "description of entity")

# example, create otherEntity object from a png file
desert_fertilization_sites <- create_otherEntity(
  targetFile = "~/Desktop/desert_fertilization_sampling_sites.png",
  description = "map of approximate location of desert fertiliztion long-term study sites")

# example, create otherEntity object from a pdf file without project naming
pass_codebook_2011 <- create_otherEntity(
  targetFile = "PASS-2011-Codebook-Feb2016rev.pdf",
  description = "PASS 2011 survey codebook",
  projectNaming = FALSE)

# example, create otherEntity object from a directory
pass_codebook_2011 <- create_otherEntity(
  targetFile = "~/Desktop/max_temperature",
  description = "rasters of max temperature years 2000-2016")

```


### spatialVector

Given a spatial vector entity in the R envionment:

`write_attributes(vector_data_entity)` will generate a template as a csv file in
the working directory based on properties of the vector_data_entity such that
metadata properties (e.g., attributeDefinition) can be added via a editor or
spreadsheet application.

`create_spatialVector(data_entity)` performs many services:

* the spatial vector data entity is written to file as a kml in the
working/project directory with the file name: *projectid_data-entity-name.kml*
or *data-entity.kml* if projectNaming is set to FALSE. (note all output is as
kml with the epsg projection 4326)
* metadata provided in the attributes templates are ingested
* a EML object of type spatialVector is returned
* note that the spatial vector data entity name should be used consistently
within the chunk, and the resulting vector_data_entity entity should have the
name: *data_entity_SV*

The `create_spatialVector` function accepts five arguments:

1. `svname`: spatial data entity name (required)
2. `description`: a description of the data object (required)
4. `baseURL`: is the base path of the web-accessible location of the data file;
the name of the resulting file will be passed to the base path to generate a
web-resolvable file path.
5. `projectNaming`: a logical indicating whether to use the project naming
convention of project-id + object name + hash of object + object file extention.
This is the default, if set to FALSE, the file/object name is unchanged and
included in the EML.


### spatialRaster

Please see the
[create_spatialRaster](https://caplter.github.io/capeml/articles/create_spatialRaster.html)
vignette for more information about creating EML for raster data.
