---
output: github_document
---

```{r libraries, echo=FALSE}
library(kableExtra)
```

<!-- README.md is generated from README.Rmd. Please edit the latter. -->

# capeml: tools to aid the generation of EML metadata

This package contains tools to aid the generation of EML metadata with intent to publish a dataset (data + metadata) in the Environmental Data Initiative (EDI) data repository. Functions and a workflow are included that allow for the creation of metadata at the data set level, and individual data entities (spatial vectors, raster, other entities, data tables). 

Helper functions for the creation of spatialRaster entities, and to generate a keywordSet from a template file are currently supported. Additional helper functions are planned.

## Installation

Install the current version from GitHub (after installing the `devtools` package from CRAN):

```{r eval=FALSE}
devtools::install_github("CAPLTER/capeml")
```


## raster data

Please see the [create_spatialRaster](https://github.com/CAPLTER/capeml/blob/master/vignettes/create_spatialRaster.Rmd) vignette for more information about creating EML for raster data.


## generate keywordSet from file

The function `create_keywordSet` geneerates a EML object of type keywordSet from a csv file containing thesaurus, keyword, and type where type is an optional keyword attribute. Keyword files should be structured like the following...

```{r sampleKeywordFile, echo=FALSE}
load(file = './data/keywordTest.rda')
kable(keywordTest)
```

Call the function and assign output:

```{r create_keywordSet, eval=FALSE}

datasetKeyords <- create_keywordSet('path/keywordFile.csv')
```

The object (e.g., datasetKeywords) can then be included in the EML generation script.
