---
title: "create_spatialRaster"
author: "S. Earl"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## detail raster metdata

For each raster, generate a separate metadata file to detail critical information about the raster. Note that the raster file name should be provided as unquoted text, and the file (or files) should be in the directory specified when calling the create_spatialRaster function.

```{r, echo=FALSE}
load(file = "CAP1985_metadata.rda")
```


```{r, echo=FALSE, results='asis'}
knitr::kable(CAP1985_metadata)
```



### categorical raster values

If the raster values are categorical, generate a metadata file to catalog the unique raster value categories and their meaning. Additional metadata about the categories can be provided at the data set level.

```{r, echo=FALSE}
load(file = "CAP1985_factors.rda")
```

```{r, echo=FALSE, results='asis'}
knitr::kable(CAP1985_factors)
```


## call the function

Call the create_spatialRaster function to generate the EML to describe the raster. Arguments include the quoted path to the raster file (or files), the quoted path and name of the raster metadata file, and the quoted path and name of the raster value categories (if needed).

```{r, eval=FALSE}
cap1985 <- create_spatialRaster('~/knb-lter-cap.650.1/CAP 30m Landsat Series Submit/',
                                '~/knb-lter-cap.650.1/rasterMetadataFiles/CAP1985_metadata.csv',
                                '~/knb-lter-cap.650.1/rasterMetadataFiles/landsat_factors.csv')
```

